Collocated Snapshot Backup with VM-Centric Deduplication in the Cloud

### 2013-09-26 snapshot_system/fast14_evaluation.tex
revise
figure updates

\section{Evaluation}
\label{sect:evaluation}
We have implemented and evaluated a prototype of our VC scheme on a Linux cluster of machines with
8-core 3.1Ghz AMD FX-8120 and 16 GB RAM. 
Our implementation is based on Alibaba cloud platform~\cite{Aliyun,WeiZhangIEEE}
and the underlying DFS uses  QFS with default replication degree 3 while the PDS replication degree is 6.
Our evaluation objective is to
study the benefit in fault tolerance and   deduplication efficiency of VC and compare it
with an alternative VO design. We also and assess backup throughput and  resource usage for a large number of VMs.

@@

\subsection{Settings}
We have performed a trace-driven study using a production dataset~\cite{WeiZhangIEEE} from 
Alibaba Aliyun's cloud platform with about 100 machines. 
Each machine hosts up to 25 VMs and each VM keeps 10 automatically-generated snapshots in the storage system while
a user may instruct extra snapshots to be saved.
The VMs of the sampled data set use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, win2008 and win2003. 
The backup of VM snapshots is required to complete within few  hours every night.
Based on our study of production  data,  each VM has about 40GB of storage  data  on average
including OS and user data disk.
%All data are divided into 2 MB fix-sized segments and 
The signature for variable-sized blocks is computed using their SHA-1 hash~\cite{similar94,rabin81}. 
%Each segment is divided into 
%variable-sized content chunks with an average size of 4KB.
%Popularity of data blocks are collected through global counting and 
%and top popular ones ($\delta=2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:store}.

\subsection{Impact on Fault Isolation}

@@

Although the loss of a PDS block affects many VMs,
by increasing replication for those blocks we minimize the effect on VM availability.
%By having extra replicas for the PDS data, we reduce the failure rate of PDS
%FSBs enough that the failure rate for the VM approaches the failure rate of the
%Non-PDS blocks. This minimizes the drop in availability between the expected
%and worst case for VC.
%
%One of the advantages of VC is that there is a division between
%popular and non-popular data to facilitate improved reliability. 
Figure~\ref{fig:pds-replication} shows
the impact of increasing PDS data replication. 
While the impact on storage cost is small, 
a replication degree of 6  has a significant improvement over 4, but the availability is about
the same for $r_c=6$ and $r_c=9$ (beyond $r_c=6$ improvements are minimal).

@@

                    ylabel={\tiny VM Availability (\%)},

@@

\end{figure}
}
\begin{table}
  \centering
    \footnotesize
    \tabcolsep=0.11cm
    \pgfplotstabletypeset[
        columns={NodesFailed,VO100,VC100,VO1000,VC1000},
        columns/NodesFailed/.style={
            column name={\multirow{3}{*}{Failures ($d$)}}
        },
        columns/VO100/.style={
            column name={
                \multicolumn{4}{c|}{VM Availability(\%)}\\&
                \multicolumn{2}{c|}{$p=100$}&\multicolumn{2}{c|}{$p=1000$}\\&
            VO},
            fixed, precision=6
        },
        columns/VC100/.style={
            column name={VC},
            fixed, precision=6},
        columns/VO1000/.style={
            column name={VO},
            fixed, precision=6},
        columns/VC1000/.style={
            column name={VC},
            fixed, precision=6},
        every head row/.style={
            before row={\hline},
            after row={\hline},
        },
        every last row/.style={after row=\hline},
        column type/.add={}{|},
        every first column/.style={column type/.add={|}{}},
        %the below rows can be used to extract rows 3,5,10,20 from rows 0-40
        %skip rows between index={0}{3},
        %skip rows between index={4}{5},
        %skip rows between index={6}{10},
        %skip rows between index={11}{20},
        %skip rows between index={21}{41}
    ]{figures/availability_table.txt}
    \caption{Availability of VM snapshots for VO and VC.}
    \label{tab:vm-availability}
\end{table}

@@

Figure~\ref{fig:exbin-efficiency-graph2} shows the deduplication efficiency for SRB and VC.
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
Deduplication efficiency is defined as the percent of duplicate chunks
which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
 We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB per machine
and  the deduplication efficiency can reach over 96.33\%.
When $\sigma=4\%$, the deduplication efficiency can reach 96.9\% while space consumption increases to 200MB per machine. 
%and  the deduplication efficiency can reach 96.489\%.
%When $\sigma=4\%$, the deduplication efficiency can reach 97.027\% while space consumption increases to 200MB per machine. 
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster to support fast in-memory PDS index lookup. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB can deliver up to 97.79\% deduplication efficiency which is slightly better than VC.
Thus this represents a tradeoff that VC provides better fault tolerance and fast approximate deletion
with competitive deduplication efficiency.

@@

Figure~\ref{fig:exbin-efficiency-graph2} also shows the curve  of VC without local similarity search.
Adding local similarity-guided search increases deduplication efficiency from 93\% to over 96\%.
That is one significant improvement compared to the work in ~\cite{WeiZhangIEEE}
which uses the parent segment at the same offset to detect duplicates locally.
The reason is that  
%Without this (i.e. when we only 
when a data  segment is moved to another location on disk, for example when a file is rewritten
rather than modified in place,  dirty-bit based detection would not be able to detect such a movement.
 We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently,
so the similarity search becomes especially important for those VMs.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.
In general, our experiments show that
dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
which leads to about a 75.86\% reduction.
Similarity-guided local search can further reduce the data size
to about 12.05\% of original, namely it delivers a 50.08\% reduction to the dirty segments.
The popularity-guided global deduplication with $\sigma=2\% $
can reduce the data further to 8.6\% of its original size, so
it provides additional 28.63\% reduction to the remaining data.

@@

{\bf Storage cost of replication.} 
%Table ~\ref{tab:replication_cost} shows the total storage space required by 
%VC as the PDS replication degree changes while non-PDS replication degree is fixed as 3. 
The total storage  for all VM snapshots in each physical machine takes about 3.065TB on average when
the replication degree of both PDS and non-PDS data is 3. 
Allocating  one extra  copy for PDS data only adds  7GB in total per machine.
Thus PDS replication degree 6 only increases the total space by 0.685\% while PDS replication degree 9 adds 1.37\% 
space overhead, which is still small.
%The increase in storage cost is minimal because the PDS makes up a 
%small percent of the total storage space, while increasing replication degree has a  more significant benefit for
%availability as shown in Figure~\ref{fig:pds-replication}.

\begin{table}
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline
    Tasks & CPU & Mem &Read &  Write  & Time  \\ 
    & & (MB)          &(MB/s) &  (MB/s) & (hrs) \\ \hline
%    1     & 19\% & 118.1 & 50MB/s 8.37 & 1.314\\ \hline
    1     & 19\% & 118 & 50 & 16.4 & 1.31\\ \hline
%    2     & 35\% & 131.8 & 9.0 & 1.226\\ \hline
    2     & 35\% & 132 &50  & 17.6 & 1.23\\ \hline
    %4     & 63\% & 154.1 & 9.3 & 1.179\\ \hline
    4     & 63\% & 154&50   & 18.3 & 1.18\\ \hline
%    6     & 77\% & 171.9 & 9.46 & 1.162\\ \hline
%    8     & 82\% & 90.5 & 89.2 & \\ \hline
%    10                      & 85\% & 97.2   & 90.4 \\ \hline
%    12                      & 91\% & 95.6    & 91.5 \\ \hline
    \end{tabular}
\caption{Resource usage of concurrent backup tasks at each machine}
\label{tab:resource_usage}
\end{table}

{\bf Memory and disk bandwidth usage with multi-VM processing}. 
We have further studied the memory and disk bandwidth usage 
when running concurrent VM snapshot backup on each machine with $\sigma=2\%$. 
Table ~\ref{tab:resource_usage} gives the resource usage  when
running 1 or multiple  VM backup tasks at the same time on each physical machine. 
``CPU'' column is the percentage of a single core used. 
``Mem'' column includes 100MB memory usage for PDS index and other space cost for executing deduplication tasks such as 
receipt metadata and cache. 
``Read'' column is controlled as 50MB/s bandwidth usage with I/O throttling so that other cloud services are not impacted too much.
``Write'' column is the I/O write usage of QFS and notice that each QFS write leads writes in multiple machines due to
data replication.     
% [ NOT REASONABLE: Notice that memory and CPU used by PDS index and QFS are not included since they belong to 
% the cloud infrastructure services and are shared among many applications.]
%We see our hybrid deduplication scheme only occupies a small amount of system resources. 
%The  local deduplication only needs to keep the parent snapshot recipe and a few similar segment recipes in 
%memory during duplicate detection.

Table ~\ref{tab:resource_usage} shows that  a single backup task per node can complete the backup of the entire
VM cluster in about 1.58 hours. Since there are about 25 VMs per machine, we could execute
more tasks in parallel at each machine. But adding more backup concurrency does not shorten the overall time
significantly because of the controlled  disk read time.
%,  we can still process raw data at near 175 MB/s. If we consider the extreme case in which each machine has 25 VMs
%at 40GB size, our snapshot system can finish backing up all the VMs (1TB total) in 1.58 hours.

{\bf Processing Time breakdown}.
Figure~\ref{fig:vc_srb_combined} shows
the  average  processing  time of  a VM segment under VC and SRB. 
VC uses $\sigma=2\%$ and $4\%$.
It has a breakdown of time for reading data and updating the metadata, network latency to visit
a remote machine, and index access for fingerprint comparison.
The change of $\sigma$ does not significantly affect the overall backup speed as
PDS lookup takes only a small amount of time.
SRB has higher index access and fingerprint comparison costs because all chunk signatures for whom the current
machine does not control the index must be sent over the network, and then rely
on the index holding machine to access its index (often on the disk, supported by a Bloom filter) 
and perform comparison.
VC is faster in this regard because it conducts in-memory local search first and then
accesses  the PDS on distributed shared memory, so most chunks can be dealt with locally.  
SRB spends  slighter more time in  reading data and updates because it also updates the on-disk
local meta data index in addition to partition indices.
Overall,  VC is faster than SRB, though data reading dominates the processing time for both algorithms.
%Figure~\ref{fig:vc_srb_combined} also reports the average backup time for a VM in VC when
%varying the PDS size.  It lists the time distribution for data reading,
%similarity-guided local search, global PDS lookup, and non-duplicate data writing. 
%While data reading dominates the backup process, PDS lookup spends about a similar amount
%of time as local search.

>>>>

\section{Evaluation}
\label{sect:evaluation}
We have implemented and evaluated a prototype of our VC scheme on a Linux cluster of machines with
8-core 3.1Ghz AMD FX-8120 and 16 GB RAM. 
Our implementation is based on Alibaba cloud platform~\cite{Aliyun,WeiZhangIEEE}
and the underlying DFS uses  QFS with default replication degree 3 while the PDS replication degree is 6.
Our evaluation objective is to
study the benefit in fault tolerance and   deduplication efficiency of VC,  
and assess its backup throughput and  resource usage.

@@

\subsection{Settings}
We have performed a trace-driven study using a production dataset~\cite{WeiZhangIEEE} from 
Alibaba Aliyun's cloud platform with about 100 machines. 
Each machine hosts up to 25 VMs and each VM keeps 10 automatically-generated snapshots in the storage system while
a user may instruct extra snapshots to be saved.
The VMs of the sampled data set use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, win2008 and win2003. 
%The backup of VM snapshots is required to complete within few  hours every night.
Based on our study of production  data,  each VM has about 40GB of storage  data  on average
including OS and user data disk.
%All data are divided into 2 MB fix-sized segments and 
The fingerprint for variable-sized chunks is computed using their SHA-1 hash~\cite{similar94,rabin81}. 
%Each segment is divided into 
%variable-sized content chunks with an average size of 4KB.
%Popularity of data blocks are collected through global counting and 
%and top popular ones ($\delta=2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:store}.

\subsection{Fault Isolation and Snapshot Availability}

@@

Although the loss of a PDS block affects many VMs,
by increasing replication for those blocks we minimize the effect on snapshot availability.
%By having extra replicas for the PDS data, we reduce the failure rate of PDS
%FSBs enough that the failure rate for the VM approaches the failure rate of the
%Non-PDS blocks. This minimizes the drop in availability between the expected
%and worst case for VC.
%
%One of the advantages of VC is that there is a division between
%popular and non-popular data to facilitate improved reliability. 
Figure~\ref{fig:pds-replication} shows
the impact of increasing PDS data replication. 
While the impact on storage cost is small, 
a replication degree of 6  has a significant improvement over 4, but the availability is about
the same for $r_c=6$ and $r_c=9$ (beyond $r_c=6$ improvements are minimal).

@@

                    ylabel={\tiny Snapshot Availability (\%)},

@@

\end{figure}
}
\begin{table}
  \centering
    \footnotesize
    \tabcolsep=0.11cm
    \pgfplotstabletypeset[
        columns={NodesFailed,VO100,VC100,VO1000,VC1000},
        columns/NodesFailed/.style={
            column name={\multirow{3}{*}{Failures ($d$)}}
        },
        columns/VO100/.style={
            column name={
                \multicolumn{4}{c|}{Snapshot Availability(\%)}\\&
                \multicolumn{2}{c|}{$p=100$}&\multicolumn{2}{c|}{$p=1000$}\\&
            VO},
            fixed, precision=6
        },
        columns/VC100/.style={
            column name={VC},
            fixed, precision=6},
        columns/VO1000/.style={
            column name={VO},
            fixed, precision=6},
        columns/VC1000/.style={
            column name={VC},
            fixed, precision=6},
        every head row/.style={
            before row={\hline},
            after row={\hline},
        },
        every last row/.style={after row=\hline},
        column type/.add={}{|},
        every first column/.style={column type/.add={|}{}},
        %the below rows can be used to extract rows 3,5,10,20 from rows 0-40
        %skip rows between index={0}{3},
        %skip rows between index={4}{5},
        %skip rows between index={6}{10},
        %skip rows between index={11}{20},
        %skip rows between index={21}{41}
    ]{figures/availability_table.txt}
    \caption{Availability of VM snapshots for VO and VC.}
    \label{tab:vm-availability}
\end{table}

@@

Figure~\ref{fig:exbin-efficiency-graph2} shows the deduplication efficiency for SRB and VC,
% <<<<<<< HEAD
% Our deduplication efficiency is defined as the percent of duplicate chunks
% which are detected and deduplicated, so only global perfect deduplication can have 100\% efficiency.
% The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
% is defined in Table~\ref{tab:symbol}. 
% =======
namely the percent of duplicate chunks
which are detected and removed.
% so only global perfect deduplication can have 100\% efficiency.
% We also compare several PDS sizes chosen for VC.
%(see Table~\ref{tab:symbol} for definition of ``$\sigma=2\%$'')
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB per machine
and  the deduplication efficiency can reach over 96.33\%.
When $\sigma=4\%$, the deduplication efficiency can reach 96.9\% while space consumption increases to 200MB per machine. 
%and  the deduplication efficiency can reach 96.489\%.
%When $\sigma=4\%$, the deduplication efficiency can reach 97.027\% while space consumption increases to 200MB per machine. 
The loss of efficiency in VC is caused by the restriction of the physical memory available
in the cluster for fast in-memory PDS index lookup. 
%In the extreme case if we put all the fingerprint into PDS 
%index ($\sigma=100\%$) then VC will have 100\% efficiency.
SRB can deliver up to 97.79\% deduplication efficiency which is slightly better than VC.
Thus this represents a tradeoff that VC provides better fault tolerance and fast approximate deletion
with competitive deduplication efficiency.

@@

Figure~\ref{fig:exbin-efficiency-graph2} also shows the curve  of VC without local similarity search.
There is a big efficiency drop in this curve when the number of VMs is about 30.
The reason is that  
%Without this (i.e. when we only 
when a data  segment is moved to another location on disk, for example when a file is rewritten
rather than modified in place,  
a dirty-bit or offset based detection would not be able to detect such a movement and 
similarity search becomes especially important.  We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently.
In general, adding local similarity-guided search increases deduplication efficiency from 93\% to over 96\%.
That is one significant improvement compared to the work in ~\cite{WeiZhangIEEE}
which uses the parent segment at the same offset to detect duplicates instead of similarity-guided search.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.

In general, our experiments show that
dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
which leads to about a 75.86\% reduction.
Similarity-guided local search can further reduce the data size
to about 12.05\% of original, namely it delivers a 50.08\% reduction to the dirty segments.
The popularity-guided global deduplication with $\sigma=2\% $
can reduce the data further to 8.6\% of its original size, so
it provides additional 28.63\% reduction to the remaining data.

@@

{\bf Storage cost of replication.} 
%Table ~\ref{tab:replication_cost} shows the total storage space required by 
%VC as the PDS replication degree changes while non-PDS replication degree is fixed as 3. 
When the replication degree of both PDS and non-PDS data is 3, 
the total storage  for all VM snapshots in each physical machine takes about 3.065TB on average before compression
and 0.75TB after compression.  Allocating  one extra  copy for PDS data only adds  7GB in total per machine.
Thus PDS replication degree 6 only increases the total space by 0.685\% while PDS replication degree 9 adds 1.37\% 
space overhead, which is still small.
%The increase in storage cost is minimal because the PDS makes up a 
%small percent of the total storage space, while increasing replication degree has a  more significant benefit for
%availability as shown in Figure~\ref{fig:pds-replication}.

\begin{table}
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline
    Tasks & CPU & Mem &Read &  Write  & Time  \\ 
    & & (MB)          &(MB/s) &  (MB/s) & (hrs) \\ \hline
%    1     & 19\% & 118.1 & 50MB/s 8.37 & 1.314\\ \hline
    1     & 19\% & 118 & 50 & 16.4 & 1.31\\ \hline
%    2     & 35\% & 131.8 & 9.0 & 1.226\\ \hline
    2     & 35\% & 132 &50  & 17.6 & 1.23\\ \hline
    %4     & 63\% & 154.1 & 9.3 & 1.179\\ \hline
    4     & 63\% & 154&50   & 18.3 & 1.18\\ \hline
    6     & 77\% & 171.9 & 9.46 & 1.162\\ \hline
%    8     & 82\% & 90.5 & 89.2 & \\ \hline
%    10                      & 85\% & 97.2   & 90.4 \\ \hline
%    12                      & 91\% & 95.6    & 91.5 \\ \hline
    \end{tabular}
\caption{Resource usage of concurrent backup tasks at each machine}
\label{tab:resource_usage}
\end{table}

{\bf Memory and disk bandwidth usage with multi-VM processing}. 
We have further studied the memory and disk bandwidth usage 
when running concurrent VM snapshot backup on each machine with $\sigma=2\%$. 
Table ~\ref{tab:resource_usage} gives the resource usage  when
running 1 or multiple  VM backup tasks at the same time on each physical machine. 
``CPU'' column is the percentage of a single core used. 
``Mem'' column includes 100MB memory usage for PDS index and other space cost for executing deduplication tasks such as 
receipt metadata and cache. 
``Read'' column is controlled as 50MB/s bandwidth usage with I/O throttling so that other cloud services are not impacted too much.
The peak read performance is about 300MB and we only use 16.7\% with this collocation consideration.
``Write'' column is the I/O write usage of QFS and notice that each QFS write leads writes in multiple machines due to
data replication.     50MB dirty segment read speed triggers about 16.4MB/second disk write for non duplicates with one backup task.
% [ NOT REASONABLE: Notice that memory and CPU used by PDS index and QFS are not included since they belong to 
% the cloud infrastructure services and are shared among many applications.]
%We see our hybrid deduplication scheme only occupies a small amount of system resources. 
%The  local deduplication only needs to keep the parent snapshot recipe and a few similar segment recipes in 
%memory during duplicate detection.

Table ~\ref{tab:resource_usage} shows that  a single backup task per node can complete the backup of the entire
VM cluster in about 1.31 hours. Since there are about 25 VMs per machine, we could execute
more tasks in parallel at each machine. But adding more backup concurrency does not shorten the overall time
significantly because of the controlled  disk read time.
%,  we can still process raw data at near 175 MB/s. If we consider the extreme case in which each machine has 25 VMs
%at 40GB size, our snapshot system can finish backing up all the VMs (1TB total) in 1.58 hours.

{\bf Processing Time breakdown}.
Figure~\ref{fig:vc_srb_combined} shows
the  average  processing  time of  a VM segment under VC and SRB. 
VC uses $\sigma=2\%$ and $4\%$.
It has a breakdown of processing time.
``Data read/write'' includes disk reading and writing, and 
updating of the metadata.
`` Network cost" includes the cost of transferring snapshot and meta data from one machine to another during backup.
``Index access/lookup'' includes the time for 
latency to visit
a remote machine, and index access for fingerprint comparison.
The change of $\sigma$ does not significantly affect the overall backup speed as
PDS lookup takes only a small amount of time.
SRB has higher index access and fingerprint comparison costs because all chunk signatures for whom the current
machine does not control the index must be sent over the network, and then rely
on the index holding machine to access its index (often on the disk, supported by a Bloom filter) 
and perform comparison.
VC is faster in this regard because it conducts in-memory local search first and then
accesses  the PDS on distributed shared memory, so most chunks can be dealt with locally.  
SRB spends  slighter more time in  reading data and updates because it also updates the on-disk
local meta data index in addition to partition indices.
Overall,  VC is faster than SRB, though data reading dominates the processing time for both algorithms.
%Figure~\ref{fig:vc_srb_combined} also reports the average backup time for a VM in VC when
%varying the PDS size.  It lists the time distribution for data reading,
%similarity-guided local search, global PDS lookup, and non-duplicate data writing. 
%While data reading dominates the backup process, PDS lookup spends about a similar amount
%of time as local search.
